---
title: "STA 522 HW6 (Multi-Stage Sampling)"
author: "Daniel Truver"
date: "2/21/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### (1) Lohr, Chapter 6 Problem 4  

We begin with showing $\hat{t}_\psi$ is unbiased. 
$$
E(\hat{t}_\psi) 
= E\left( \frac{t_i}{\psi_i} \right) 
= \sum_{i} \left( \frac{t_i}{\psi_i} \right) \psi_i
= \sum_i t_i = t.
$$

```{r 1.lazyCalculation, include=FALSE}
psi = c(7/16, rep(3/16,3))
t = c(11,20,24,245)
variance = 0
for (i in 1:4){
  variance = variance + (t[i]/psi[i] - 300)^2 * psi[i]
}
```

The variacne of $\hat{t}_\psi$ is given by the following.
$$
\begin{aligned}
Var(\hat{t}_\psi) 
&= E\left[(\hat{t}_\psi - t)^2\right] \\
&= \sum_i \left( \frac{t_i}{\psi_i} - t \right)^2 \psi_i \\
&= (11(16/7) - 300)^2(7/16) + (20(16/3) - 300)^2(3/16)\\
&\quad\quad + (24(16/3) - 300)^2(3/16) + (245(16/3) - 300)^2(3/16) \\
&\approx 2.36 \times 10^5
\end{aligned}
$$

This variance is much worse than the variance of the $\pi$ps estimator used in section 6.1, an estimator which is also unbiased. Therefore, we do not have a reason to use this distribution for $\psi_i$. Intuitively, it does not make sense to put so much weight on the smallest of the supermarkets. 

#### (2) Lohr, Chapter 6 Problem 12

Data from example 6.5 

#### (3) Lohr, Chapter 6 Problem 13

Data from example 6.5

